---
title: "hw1math4323"
author: "anthonycastillo   UH ID: 1670011"
date: "2022-09-03"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE}
Credit <- read.csv("C:/Users/casti/Downloads/Credit.csv", stringsAsFactors=TRUE)
```

Question 1a:

classification -> prediction n=20 p=14

Question 1b:

regression->inference n=200 p=15

Question 1c:

regression->inference n = 1500 p=10
Question 1d:

classification->prediction n = 300 p =8

Question 2a:

supervised learning

Question 2b:

i. for example (a) and (d) f_hat can be treated as a black box. The other examples (b) and (c) they can not be treated as a black box.

ii.for inference examples we need to know the exact form of f_hat. The predication example we can use the treat as black box because the form of f_hat is not very important as long the prediction is accurate.


Question 2c:

reducible and irreducible errors. The reducible error can be improved using a better statistical learning technique. The irreducible error can not be improved because we can not avoid randoms errors in the model.


Question 3:

example 1: streaming services trying to group similar tv show/movies that the consumer has watched

example 2: selling a product that targets a specific audience like video games

example 3: in sports you can group together many differnet statistics to help the team win/lose the game 

Question 4a:

Use RStudio’s drop-down menu (Environment → Import Dataset → From Text (base) ...)
to read the data into R. Make sure the Heading is set to Y es. Call the loaded
data Credit.


Question 4b-1:
```{r}
summary(Credit)
```

Question 4b-2:
numerical values - Income, Limit, Rating, Cards, Age, Education, Balance
 
categorical values - Own, Student, Married, Region
 
Question 4b-3:
```{r}
pairs(~Income+Limit+Rating+Cards+Age+Education+Balance, Credit)
```


Question 4b-4:
```{r}
plot(Credit$Student,Credit$Balance)
```
 
 
 Question 4b-5:
```{r}
high=(Credit$Rating>680)
table(high)
#plot(Credit$Balance,Credit$high)
```
 
 
 Question 4b-6:
```{r}
par(mfrow = c(2,2))
hist(Credit$Income)
hist(Credit$Age)
hist(Credit$Balance)
hist(Credit$Rating)
```
 
 
 Question 5a:
 
```{r}
library(MASS)
#Boston
?Boston
```
 
 506 rows and 14 columns. 
 lstat - lower status of the population
 ptratio - pupil-teacher ratio by town
 chas - if bounded by charles river or not
 medv - median value of owner-occupied homes in $1000s
 
 Question 5b:
 
 most of the predictors are quantitative variables
 
 Question 5c:
 
```{r}
Boston_range = sapply(Boston,range)
rownames(Boston_range) = c("min:","max:")
Boston_range
```
 
 
 Question 5d:
```{r}
Boston_mean_sd = sapply(Boston, function(x) c(mean(x),sd(x)))
rownames(Boston_mean_sd) = c("mean:","sd:")
Boston_mean_sd
```
 
 Question 5e:
 
```{r}
newBoston = Boston[-c(50:100),]
NB_range_mean_sd = sapply(newBoston, function(x) round(c(range(x),mean(x),sd(x)),digits=1))
#format(round(NB_range_mean_sd,digits = 4),nsmall = 4)
rownames(NB_range_mean_sd) = c("min:","max:","mean:", "sd:")
NB_range_mean_sd
```
 
 Question 5f:
 
```{r}
#pairs(Boston)
par(mfrow=c(2,2))
plot(Boston$medv,Boston$crim)
plot(Boston$age,Boston$crim)
plot(Boston$lstat,Boston$crim)
plot(Boston$zn,Boston$crim)

```
 crim and medv - as medv increases the crim rate decreases.
 
 crim and lstat - as lstat increases there is somewhat of an increase in crime rates.
 
 crim and age - as age increase the crim rate also increases.
 
 crim and zn - 0 zn have higher crim rate than the rest of the plot.
 
 
 Question 5g:
```{r}
par(mfrow=c(2,2))
hist(Boston$crim)
hist(Boston$tax)
hist(Boston$ptratio)
```
 
 
 Question 5h:
```{r}
nrow(subset(Boston,chas ==1))
```
 
 
 
 Question 5i:
 
```{r}
#pairs(Boston)
par(mfrow=c(2,2))
plot(Boston$crim, Boston$medv)
plot(Boston$lstat, Boston$medv)
plot(Boston$black, Boston$medv)
plot(Boston$rm, Boston$medv)
```
 
 
