---
title: "HW5-MATH4323"
author: "anthonycastillo ID:1670011"
date: "2022-11-01"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1 (a):\
```{r}
x1 <- c(1,2,3,6,6,7)
x2 <- c(4,2,3,2,1,0)
plot(x1,x2, xlab = "x1", ylab = "x2", pch = 19, col = "blue")
```

Question 1 (b):\
```{r}
RNGkind(sample.kind = "default")
set.seed(2)
labels <- sample(2,6, replace = T)
labels
plot(x1,x2, pch = 19, col = labels ,xlab = "x1",ylab = "x2")
```

Question 1 (c):\
```{r}
cluster1x <- sum(x1[1:2])/2
cluster1y <- sum(x2[1:2])/2
cluster2x <- sum(x1[3:6])/4
cluster2y <- sum(x2[3:6])/4
```
Question 1 (d):\
```{r}
clusterassign <- c(1,1,1,2,2,2)
plot(x1,x2, pch = 19, col = clusterassign,xlab = "x1",ylab = "x2")
points(cluster1x,cluster1y, col = 1, pch = 4)
points(cluster2x,cluster2y, col = 2, pch = 4)
```

assign centroid (1.5,3) to cluster 1\
assign centroid (5.5,1.5) to cluster 2\
Question 1 (e):\
```{r}
up.cluster1x <- sum(x1[1:3])/3
up.cluster1y <- sum(x2[1:3])/3
up.cluster2x <- sum(x1[4:6])/3
up.cluster2y <- sum(x2[4:6])/3
```
The observations stayed the same after one iteration \

Question 1 (f):\
```{r}
clusterlabel <- c(1,1,1,2,2,2)
plot(x1,x2, pch = 19, col = clusterlabel, xlab = "x1",ylab = "x2")
points(up.cluster1x,up.cluster1y, col = 1, pch = 4)
points(up.cluster2x,up.cluster2y, col = 2, pch = 4)
```

Question 2 (a):\
```{r}
library(MASS)
newBoston <- Boston[,-4]
```
Question 2 (b):\
```{r}
newBoston <- newBoston[,-13]
```
Question 2 (c):\
```{r}
boston.pca <- prcomp(newBoston, scale = TRUE)
#summary(boston.pca)
boston.pca$sdev
```
Question 2 (d):\
```{r}
boston.var <- boston.pca$sdev^2
print(boston.var)
```
Question 2 (e):\

Question 3 (a):\
```{r}
set.seed(1)
x<- matrix(rnorm(4*5*2),ncol=2)

#split in 5 clusters
x[1:4,2] <- x[1:4,2]+6
x[5:8,1] <- x[5:8,1]+9
x[9:12,1] <- x[9:12,1]-5
x[13:16,2] <- x[13:16,2]-5
scale(x)
plot(x, pch=19, main = "Simulated Data",
     ylab = "x2", xlab = "x1")
```

Question 3 (b):\
```{r}
x <- data.frame(x)
library(factoextra)
sim.wss <- numeric(6)
for(i in 4:6){
  sim.wss[i] <- eclust(x,
                    FUNcluster = "kmeans",
                    k=i,
                    nstart=50,
                    graph = 0)$tot.withinss
}
print(sim.wss[4:6])
```

going from k=4 to k=5 had a bigger drop in total WSS\
Question 3 (c):\
```{r}
for(i in 4:6){
  eclust(x,FUNcluster = "kmeans",
         k=i,
         nstart = 50)
}
```

The clusters got smaller when increasing the K-value meaning the total distance between each value in each cluster got smaller\
Question 3 (d):\
```{r}
k.max <- 10
sim.wss <- numeric(k.max)
for(i in 1:k.max){
  sim.wss[i] <- eclust(x,
                       FUNcluster = "kmeans",
                       k = i,
                       nstart = 50,
                       graph = 0)$tot.withinss
}
plot(sim.wss, pch = 20, type = "b",
     main = "WSS for k=1....10",
     ylab = "Total WSS",
     xlab = "number of clusters K")
```

The optimal k value seems to be k=5\
Question 3 (e):\
```{r}
sim.gap <- fviz_nbclust(x,
                        kmeans,
                        nstart = 50,
                        nboot = 50,
                        method = "gap_stat")
plot(sim.gap)
```

The optimal k values is 5 which also match the number of separated clusters in the simulated data\
Question 4 (a):\
```{r}
newIris <- iris[-5]
k.max <- 10
iris.wss <- numeric(k.max)
for(i in 1:k.max){
  iris.wss[i] <- eclust(newIris,
                        FUNcluster = "kmeans",
                        k = i,
                        nstart = 50,
                        graph = 0)$tot.withinss
}
print(iris.wss)
plot(iris.wss, pch = 19, type = "b",
     main = "WSS for k=1....10",
     ylab = "Total WSS",
     xlab = "number of clusters K")
```

Question 4 (b):\
```{r}
iris.gap <- fviz_nbclust(newIris,
                         kmeans,
                         nstart = 50,
                         nboot = 50,
                         method = "gap_stat")
plot(iris.gap)
```

The optimal K value seems to be k=6\
Question 4 (c):\
```{r}
iris.clust <- eclust(newIris,
                       FUNcluster = "kmeans",
                       nstart = 50,
                      nboot  = 50,
                       graph = 0)
table(iris$Species[iris.clust$cluster == 6])
table(iris$Species[iris.clust$cluster == 5])
table(iris$Species[iris.clust$cluster == 4])
table(iris$Species[iris.clust$cluster == 3])
table(iris$Species[iris.clust$cluster == 2])
table(iris$Species[iris.clust$cluster == 1])
```
Question 4 (d):\
```{r}
scale(newIris)
iris.gap <- fviz_nbclust(newIris,
                         kmeans,
                         nstart = 50,
                         nboot = 50,
                         method = "gap_stat")
plot(iris.gap)
```
Question 4 (e):\
```{r}
iris.clust <- eclust(newIris,
                       FUNcluster = "kmeans",
                       nstart = 50,
                      nboot  = 50,
                       graph = 0)
table(iris$Species[iris.clust$cluster == 6])
table(iris$Species[iris.clust$cluster == 5])
table(iris$Species[iris.clust$cluster == 4])
table(iris$Species[iris.clust$cluster == 3])
table(iris$Species[iris.clust$cluster == 2])
table(iris$Species[iris.clust$cluster == 1])
```
Question 4 (f):\
```{r}
iris.optimal <- eclust(newIris,
                       FUNcluster = "kmeans",
                       k = 6,
                       nstart = 50)
```
Question 5 (a):\
```{r}
temp <- read.csv("E:/fall22/math4323/Ch10Ex11.csv", header = F)
data<-t(temp)
data <- data.frame(data)
```
Question 5 (b):\
```{r}
library(factoextra)
k.max <- 10
wss <- numeric(k.max)
for(i in 1:k.max){
  wss[i] <- eclust(data,
                  FUNcluster = "kmeans",
                  k=i,
                  nstart=50,
                  graph = 0)$tot.withinss
}
plot(wss, type = "b",
     pch = 19, col = "blue",
     main = "total within-cluster sum of squares for K=1-10",
     xlab = "Number of Clusters K",
     ylab = "Total WSS")
```

Yes there is an elbow at K=2\
Question 5 (c):\
```{r}
data.gap <- fviz_nbclust(data,
                         kmeans,
                         nstart = 50,
                         nboot = 20,
                         method = "gap_stat")
plot(data.gap)
```

The optimal k value is k=2\
Question 5 (d):\
```{r}
data.optimal <- eclust(data,
                       FUNcluster = "kmeans",
                       nboot = 20,
                       nstart = 50)
table(temp[1,data.optimal$cluster==2])
table(temp[1,data.optimal$cluster==1])
```
Yes the answer is close to part c\
Question 5 (e):\
```{r}
data.optimal <- eclust(data,
                       FUNcluster = "kmeans",
                       k=2,
                       nstart = 50)
```

Yes one cluster only contain healthy and the other contains the diseased group\